---
title: "Budget Impact Analysis of introducing Atypical antipsychotics"
author: "Joshua Edefo"
date: "2024-08-26"
email: "edefojoshua2000@yahoo.com"
output: github_document
---
Libraries

```{r s, message=FALSE}
library(ggplot2)
library(usethis)
```
Parameters:
Population Size: Mean = 2000, SD = 500
Typical Antipsychotics (old drug) Cost: Mean = NGN 1000, SD = NGN 250
Atypical Antipsychotics (new drug) Cost: Mean = NGN 6000, SD = NGN 1500
Currency: NGN (Nigerian Naira)
Time_horizon =  5 years (2016 -2020)
Yearly market share growth = 5%

```{r b}

#Set the number of simulations

n_simulations <- 10000

# Set the specific years for the time horizon
years <- c(2016, 2017, 2018, 2019, 2020)

# Define input distributions
set.seed(123)  # For reproducibility

# Yearly market share growth
market_share_growth_rate <- 1.05  # 5% growth in market share each year (you can adjust this)

# Initialize lists to store yearly budget impacts
budget_impact_per_year <- list()

for (year_index in 1:length(years)) {
  # Simulate population size for each year
  population_size <- rnorm(n_simulations, mean = 2000, sd = 500)  # Normal distribution for patient #population
  
  # Simulate market share for Drug B (increases yearly)
  market_share_B <- rbeta(n_simulations, shape1 = 3, shape2 = 7) * (market_share_growth_rate ^ (year_index - 1))
  
  # Simulate costs for Drug A (typical antipsychotics) and Drug B (atypical antipsychotics)
  cost_A <- rnorm(n_simulations, mean = 1000, sd = 250)
  cost_B <- rnorm(n_simulations, mean = 6000, sd = 1500)
  # Calculate yearly costs
  cost_old <- population_size * cost_A  # Old scenario (only Drug A)
  cost_new <- population_size * ((1 - market_share_B) * cost_A + market_share_B * cost_B)  # New scenario (Drug A and B)
  
  # Calculate yearly budget impact
  budget_impact <- cost_new - cost_old
  
  # Store the results for each year
  budget_impact_per_year[[year_index]] <- budget_impact
}
```

Budget Impact assessment and plot of the dustribution

```{r c}

# Summing the budget impact over the specified years
total_budget_impact <- Reduce("+", budget_impact_per_year)

# Plot the distribution of total budget impact over the 5 years
ggplot(data.frame(total_budget_impact), aes(x = total_budget_impact)) +
  geom_histogram(binwidth = 1000000, fill = "skyblue", color = "black") +
  labs(title = "Monte Carlo Simulation of Budget Impact (2016-2020)",
       x = "Total Budget Impact (NGN)",
       y = "Frequency") +
  theme_minimal()


```


The impact estimates

```{r d}
# Calculate and print overall key statistics for the 5-year period
mean_impact <- mean(total_budget_impact)
ci_lower <- quantile(total_budget_impact, 0.025)
ci_upper <- quantile(total_budget_impact, 0.975)

cat(sprintf("Overall Mean Budget Impact (2016-2020): NGN %.2f\n", mean_impact))
cat(sprintf("95%% Confidence Interval: NGN %.2f to NGN %.2f\n", ci_lower, ci_upper))

# Print key statistics for each year
for (year_index in 1:length(years)) {
  yearly_mean <- mean(budget_impact_per_year[[year_index]])
  yearly_sd <- sd(budget_impact_per_year[[year_index]])
  yearly_ci_lower <- quantile(budget_impact_per_year[[year_index]], 0.025)
  yearly_ci_upper <- quantile(budget_impact_per_year[[year_index]], 0.975)
  
  cat(sprintf("\nYear %d Budget Impact:\n", years[year_index]))
  cat(sprintf("  Mean: NGN %.2f\n", yearly_mean))
  cat(sprintf("  Standard Deviation: NGN %.2f\n", yearly_sd))
  cat(sprintf("  95%% Confidence Interval: NGN %.2f to NGN %.2f\n", yearly_ci_lower, yearly_ci_upper))
}



```
Results:
Year 2016 Budget Impact: Mean: NGN 2969617.09; SD: NGN 1899980.82
Year 2017 Budget Impact: Mean: NGN 3184747.76; SD: NGN 2036246.57
Year 2018 Budget Impact: Mean: NGN 3293329.26, SD: NGN 2088343.26
Year 2019 Budget Impact: Mean: NGN 3482219.34;  SD: NGN 2272126.42
Year 2020 Budget Impact: Mean: NGN 3695960.70; SD: NGN 2383510.29

  
session information

```{r z}
sessionInfo()

```